Even-point Multi-loop Unitarity and its Applications: Exponentiation, Anomalies and Evanescence

### 2023-10-16 emu_JHEP_REsubmission/responseToEditor.txt

Dear Editor,

We thank the referee for their feedback yada yada yada

We have addressed all the minor clarifications:

C1] All appearances of 'Gramm determinants' have been changed to 'Gram determinants', and the appearance of 'veirbein' has been corrected to 'vierbein'.
C2] In 2.49) 'C' has been corrected to 'C_3'
C3] In 3.14 and 3.31, we have corrected 'l^{\mu_2}' to the uncontracted 'l^{\mu_3}'
C4] Below 4.39 we have corrected the doubled word 'use use'
C5] We have changed 'present as tree-level' to 'present at tree-level'
C6] At the referees request, we have changed ' adjoint double' to 'adjoint double copy' three lines about section 5.2.2

Furthermore, we very much appreciative the thoughtful and substantive feedback from the referee, and have made the following additions to the manuscript and provide the following clarification,

Q1] We absolutely agree that two-loop four-point is a particularly simple example that permits constructibility in terms of one-loop integral convolutions. Indeed this is precisely the reason that we focus on these amplitudes as a sharp example. We have clarified by changing

"As we can see from the diagrams that contribute at two-loops four-point in eq. (3.6), even- point theories can be fully constructed from convolutions of one-loop integrals. As described earlier, we call such integrals recursively one-loop since we can iteratively apply one-loop unitarity and tensor reduction methods in order to recover the full multi-loop structure. In the next two sections, we first illustrate the recursive behavior of convolution integrals, and then use their properties to develop multi-loop tensor reduction methods that we use throughout the paper."

To

"As we can see from the diagrams that contribute in eq. (3.6), the scattering amplitudes for even-point theories at two-loop four-point can be fully constructed from convolutions of one-loop integrals. As mentioned earlier, we call such integrals recursively one-loop since we can iteratively apply one-loop unitarity and tensor reduction methods in order to recover the full multi-loop structure. We note that at general multiplicity and at higher loop order, EMU cut construction will produce integrals outside the space of recursively one-loop integrals. While it would absolutely be worthwhile to determine the EMU integral basis in generality, in this paper we focus on the nice example of two-loop four-point. To this end, in the next two sections we will illustrate the recursive behavior of one-loop convolution integrals, and then use their properties to develop multi-loop tensor reduction methods for the two-loop four-point amplitudes considered in this work."

Q2] Rather than adding a forward pointer, we have moved the definition of stA^{F^3} from 3.51 to immediately after 3.41. We appreciate the suggested change, as we believe this improves the flow of the section.

Q3] We appreciate the referee's willingness to maintain the ostrich integrals as drawn, even though we agree that it obscures the Z2 symmetry of the diagram. We have added a number a couple reminders that the integral itself is symmetric under (1<->2) exchange. First in figure 3:

"Throughout the text, ostrich integrals will be drawn with an asymmetry between the “head” and the “tail” of the ostrich, but we emphasize that the integral itself is functionally symmetric under exchange of the top two external edges."

Then below eq. 4.23:

"As noted below Fig. 3, the ostrich integral above is symmetric under 1 ↔ 2 exchange of external edges."

Q4] We did not carefully check the subleading two-loop divergence, but we agree this would be an interesting direction for future study. To leave the possibility open, we have emphasized that eq. 4.38 only captures the leading divergence at each loop order, and drawn attention to the possibility of a subleading divergence from the cross term at two-loop order:

"However, we should emphasize that eq. (4.38) only considers the leading divergence at each loop order, and does not consider subleading logarithmic terms related to the cross terms of (A1-loop/Atree)2 ∼ ε−2 + ε−1log(μ) + O(1). We leave determining whether eq. (4.38) also captures the subleading divergences as a direction of future study."

Q5] We have spelled out the result of plugging in the condition for maximal supersymmetry in D=4 -- these can be found in eq. 4.67 in D-dimensions and eq. 4.72 for D=4

Furthermore, we have emphasized that the particle content of maximal SUSY in 4.82 -- 4.85 is explicitly for a single supersymmetric vector multiplet, such that $N_\gamma = 1$):

''When one plugs in the state content for a single maximally supersymmetric vector multiplet (i.e. with normalization Nγ = 1), stated below for a variety of dimensions''

Q6] Below eq. 5.47 we provide some background on how to derive these Hilbert series. Finding generating functions in general can be challenging, but since both eq 5.46 and 5.47 iteratively compound, we they can be easily derived from appropriate multiplication of 1/(1-\alpha). In eqs. 5.48 -- 5.51 we give an explicit example of how one can derive the permutation invariant Hilbert series computed in 5.46.

In addition, we have fixed the following typos that we identified in final read through:

T1] 5.46 and 5.47 we swapped the generating functions of H(ij)(kl) and H(ijkl) -- this has been resolved. This does not change the expressions of 5.52 and 5.53.
T2] 5.23 had a misplaced '+' that has been removed
T3] Below 4.19 we have fixed the typo 'Jacboi' --> 'Jacobi'

We very much appreciate the referees incredibly thoughtful feedback.

### 2023-10-16 emu_JHEP_REsubmission/emu.tex
added subdirectory, shared EMU.nb notebook
resubmission edits

The remarkable simplicity of one-loop amplitudes is even more dramatic in a fixed spacetime dimension. In $D=4$, the kinematic projection to a basis of scalar integrals saturates at the box integral due to the appearance of Gramm determinants. Roughly speaking, $N$-gon integral coefficients of \eqn{TenRed1loop} come dressed with a factor of $\mathcal{G}_N$,
\begin{equation}
\mathcal{G}_N = \text{det} (k_i \cdot k_j)\,,
\end{equation}
where $k_i$ are the external momenta flowing into the $N$-gon integral. In fixed spacetime dimension, $D$, the momenta are $D$-component vectors and thus the Lorentz product matrix $(k_i \cdot k_j)$ must have a null space for $N>D$. As a result, $C_M \neq 0$, only when $M\leq D+1$. Furthermore, in $D=4-2\epsilon$ dimensions, any scalar pentagon integral can be rewritten up to $\mathcal{O}(\epsilon)$ in terms of five pinched scalar box integrals \cite{Bern:1993kr}. All the information of one-loop 4D amplitudes can therefore be completely determined by evaluating up to box integrals, regardless of the multiplicity. We will demonstrate in the next section how this can be used in 4D integral construction.

@@

\subsection{Even-point Effective Field Theories}
\label{subsec:EPEFTReview}
In this section we provide a catalog of the even-point effective field theories studied throughout the paper. 
Some of the theories we consider, like Volkov-Akulov and supersymmetric extensions of Born-Infeld, have natural formulations in terms of veirbeins that depend on fermionic degrees of freedom. For such theories, it is useful to distinguish between the coordinate independent actions, $S$, and the coordinate-dependent Lagrangian densities, $\mathcal{L}$,
\begin{equation}
S= \int d^D x \,\mathcal{L}\,.
\end{equation}
\paragraph{Nonlinear Sigma Model}
The simplest effective field theory that we will consider is the {nonlinear sigma model} (NLSM). The chiral NLSM Lagrangian density with diagonal symmetry group $G=SU(N_c)$ can be expressed in terms of the chiral current, $j_\mu = U^\dagger \partial_\mu U$, as follows,
\begin{equation}\label{eq:NLSMLag}
\mathcal{L}^{\text{NLSM}}= \frac{1}{2}\text{tr}[(\partial_\mu U)^\dagger (\partial ^\mu U)] = \frac{1}{2}\text{tr}\left[\frac{(\partial_\mu\pi )( \partial^\mu \pi) }{(1-f_\pi^{-2}\pi^2)^2}\right]\,,
\end{equation}
where the trace is over the color indices of the gauge group, $\pi \equiv \pi^a T^a$. On the right hand side we have applied the Cayley parameterization to the $SU(N_c)$ group elements, $U= \frac{1+\pi/f_\pi}{1-\pi/f_\pi}$, where $f_\pi$ is the dimensionful pion decay width. By construction, this model has a left-right global symmetry, $G\times G$, that is nonlinearly realized, and a linearly realized diagonal subgroup, $G$. This theory is the unique two-derivative EFT that is invariant under constant shifts of the pion field in color space:
\begin{equation}\label{eq:shiftSym}
\pi^a \rightarrow \pi^a + c^a\,.
\end{equation}
We could also consider higher derivative generalizations of the NLSM that are invariant under the shift symmetry of \eqn{eq:shiftSym}.  Such  EFTs correspond to the so-called chiral limit of {chiral perturbation theory}~\cite{Weinberg:1978kz,Gasser:1983yg,Kampf:2006yf,Kampf:2013vha}, describing massless pion scattering below the chiral symmetry breaking scale of QCD:
\begin{equation}\label{eq:XPTLag}
\mathcal{L}^{\chi\text{PT}}=\frac{1}{2}\text{tr}[(\partial_\mu U)^\dagger (\partial ^\mu U)] + \frac{\beta_1}{f_\pi^2} \text{tr}[(\partial_\mu U)^\dagger (\partial ^\mu U)]^2 + \frac{\beta_2}{f_\pi^2} \text{tr}[(\partial_\mu U)^\dagger( \partial_\nu U)]\text{tr}[(\partial^\mu U)^\dagger( \partial^\nu U)] +\cdots
\end{equation}
By construction, each of the higher derivative operators appearing a bove are invariant under the shift symmetry of \eqn{eq:shiftSym}. These higher derivative generalizations of NLSM have been studied both in the context of the $S$-matrix bootstrap \cite{Manohar:2008tc,Bellazzini:2016xrt,Guerrieri:2020bto}, and soft theorem bootstraps of Refs. \cite{Cheung:2014dqa,Cheung:2015ota,Cheung:2016drk,Cheung:2018oki,Low:2019ynd}. Given the relevance of \eqn{eq:XPTLag} in Standard Model pion scattering, the amplitudes for $G= SU(2)$ isospin pions have been computed through two-loop order \cite{Bijnens:1995yn,Bijnens:1997vq,Girlanda:1997ed}. In this work, we add to the literature by computing the two-loop amplitudes of \eqn{eq:NLSMLag} for arbitrary $SU(N_c)$ color structure. We then use these amplitudes to compute $\mathcal{N}=4$ DBIVA amplitudes via the double copy in \sect{sec:DBIvDC}.

@@

\paragraph{Volkov-Akulov Fermions and Supersymmetry}
Now we will describe the supersymmetric extensions of the Born-Infeld action given above in \eqn{biLag}. First we start by defining the theory of shift symmetric fermions first written down by Volkov, Akulov (VA)~\cite{Volkov:1973ix}. The easiest way to define this shift symmetric VA theory is by constructing a volume form in the veirbein frame out of manifestly supersymmetric 1-forms. That is, define a 1-form, $ \omega^m = e^m_\mu dx^\mu$, where the veirbeins (frame fields), $e^m_\mu$, are expressed in terms of the fermion fields, $\lambda$, as follows:
 \begin{equation}
 e^m_\mu = \delta^m_\mu + i \bar{\lambda} \Gamma^m\! \stackrel{\leftrightarrow}{\partial_\mu}\!  \lambda = \delta^m_\mu + i (\bar{\lambda} \Gamma^m \partial_\mu  \lambda - \partial_\mu \bar{\lambda} \Gamma^m   \lambda)\,.
 \end{equation}
 This frame field is manifestly invariant under the superspace shift:
\begin{equation}
\delta \lambda \rightarrow \eta \qquad  \delta \bar{\lambda} \rightarrow \bar{\eta} \qquad  \delta x^\mu \rightarrow x^\mu - i (\bar{\lambda} \,\Gamma^\mu \eta- i \bar{\eta} \,\Gamma^\mu \lambda)\,.
\end{equation}
 The Volkov-Akulov action in the veirbein frame is then simply the fermionic world-volume integral over the $D$-form \cite{Kallosh:1997aw},
 \begin{equation}
 S^{\text{VA}} = \int \omega^1 \wedge \omega^2 \wedge \cdots \wedge \omega^D\,.
 \end{equation}
 This is not dissimilar from the construction of the pure scalar DBI action above. Likewise, we can write the Lagrangian in the spacetime coordinate frame in parallel to \eqn{biLag} and \eqn{DBI}. Transforming back into spacetime coordinates yields the following Lagrangian density for self interaction VA fermions upto four-point interactions:
\begin{equation}
 \label{vaLag}
  \alpha'^{\,2}\mathcal{L}^{\text{VA}} = \text{det}(e^m_\mu) =  i\bar{\lambda}\!\stackrel{\leftrightarrow}{\slashed{\partial}}\!\lambda+ \frac{1}{2}(\bar{\lambda}\Gamma^\mu \partial^\nu \lambda)(\bar{\lambda}\Gamma_\mu \partial_\nu \lambda)+\mathcal{O}(\partial^4\lambda^6) 
  \,.
\end{equation}
Furthermore, a convenient choice of field redefinition was proposed by Komargodski and Seiberg (KS) \cite{Komargodski:2009rz,Kuzenko:2010ef}, which defines a new fermion field, $\psi = \psi(\lambda,\bar{\lambda})$, in a way that is perturbatively equivalent to the VA fermions. This construction is equivalent to a theory of nilpotent chiral superfields \cite{Rocek:1978nb,Casalbuoni:1988xh,Ferrara:2014kva}, relevant for the construction of $\alpha$-attractor models of inflation \cite{Kallosh:2013yoa}. Applying the KS nonlinear field redefinition yields the considerably simpler Lagrangian for VA theory:
 \begin{equation}
   \alpha'^{\,2}\mathcal{L}^{\text{VA}} = i\bar{\psi}\slashed{\partial}\psi +\frac{1}{2} \bar{\psi}^2 \partial^2 \psi^2 + \frac{1}{4}\psi^2\bar{\psi}^2 \partial^2 \psi^2\partial^2 \bar{\psi}^2\,.
 \end{equation}
With the world-volume formulation of VA theory in hand, we can also trivially construct the action for maximal $\mathcal{N}=1$ supersymmetric Born-Infeld theory in $D=10$ \cite{Tseytlin:1999dj}. We simply replace the flat space background of \eqn{biLag} with the fermionic background of \eqn{vaLag},
\begin{equation}
 \label{DBIVALag}
S^{\text{DBIVA}}_{\mathcal{N}=1} = \int \omega^1 \wedge\omega^2 \wedge \cdots \wedge \omega^{10} \sqrt{\text{det}(\eta_{mn}+ \alpha F_{mn})}\,.
\end{equation}
It is worth noting that the fermion-vector interaction is introduced via the {non-minimal} coupling between the field strength and the fermionic spacetime metric \cite{Bergshoeff:2013pia}. This differs from the linearly realized supersymmetry of super-Yang-Mills theory, which minimally couples the vector mode to the chiral fermion. This non-minimal coupling is a requirement of supersymmtry due to the even-point nature of abelian vector theories like Born-Infeld.

@@

\subsection{On-shell Unitarity methods}\label{sec:genU}
Above we have reviewed the known behavior of one-loop integral bases in $D=4$ spacetime dimensions. In this section we will briefly demonstrate why traditional 4D unitarity methods are so powerful -- and why we must go beyond them to accommodate the multi-loop calculations of interest here. 
\paragraph{Integrands from 4D Cuts}
Applying one-loop integral reduction in 4D tells us that any one-loop massless amplitude, for which tadpoles are scaleless, can be expressed in terms of box, triangle, and bubble diagrams, with each integral weighted by dimensionally-dependent functions of kinematics:
\begin{equation}\label{eq:4DUn}
\mathcal{A}^{\text{1-loop}}= \sum_{N=2}^4 C_N(D)I_N^D = \sum_{N=2}^4 C_N(4)I_N^D+ \mathcal{R}\,,
\end{equation}
where the scalar basis-integrals, $I_N^D$, are evaluated in $D=4-2\epsilon$, and $\mathcal{R}$ is a rational term that emerges from series expanding around $\epsilon=(4-D)/2$. To construct integral coefficients, $C_N$, we make use of the Optical Theorem for the $S$-matrix, which states that,
\begin{equation}
S^\dagger S = \mathbb{1} \quad \Rightarrow \quad 2\,\text{Im}(T) = T^\dagger T\,,
\end{equation}
where $S=\mathbb{1}+iT$. Applying this relation to \eqn{eq:4DUn} allows us to identify the imaginary part of $\mathcal{A}^{\text{1-loop}}$ with the imaginary parts of each basis integral weighted by the respective 4D coefficient, $C_N(4)$:
\begin{equation}\label{eq:UnMeth1loop}
\text{Im}(\mathcal{A}^{\text{1-loop}}) = \sum_{N=2}^4 C_N(4)\, \text{Im}(I_N^D)\,.
\end{equation}
Using the optical theorem to constrain scattering amplitudes is known as the \textit{unitarity method} \cite{UnitarityMethod, BDKUniarityReview}. In order to isolate each $C_N(4)$ appearing in \eqn{eq:4DUn}, we must place particular internal legs on-shell. That is, for each internal loop propagator that we wish to take on-shell, $1/P_i^2$, we make the following replacement inside the integral:
\begin{equation}
\frac{i}{P_i^2 + i\varepsilon} \rightarrow 2\pi \delta^{(+)}(P_i^2+i\varepsilon)\,.
\end{equation}
By replacing internal propagators with positive-root delta functions, one can isolate the imaginary parts needed to determine the integral coefficients. This approach of taking individual legs on-shell is known as the method of \textit{generalized unitarity} -- and it can be systematically applied to construct all one-loop amplitudes directly from 4D unitarity cuts \cite{Forde:2007mi}. Using generalized unitarity, the integral coefficients correspond to the following set of iterated cuts:
\begin{align}
C^{(P_iP_jP_k)}_4 &\sim \boxCut\,,
\\
C^{(P_iP_j)}_3 &\sim \triCut - \sum_{k} C^{(P_iP_jP_k)}_4\,,
\\
C^{(P_i)}_2 &\sim \bubCut- \sum_{j} C^{(P_iP_j)} - \sum_{j,k} C^{(P_iP_jP_k)}_4\,,
\end{align}
where exposed legs are summed over all internal states crossing the cut, and the labels $P_i$ are external momentum scales flowing into the specified cuts. The subtractions account for the overlapping information in the respective $N$-gon cuts. By momentum conservation, the momentum flow in the unlabeled vertex is completely determined by the other $N-1$ vertices. As a concrete example, consider the 4D cut construction of the bubble coefficient for pure Born-Infeld theory:
\begin{align}\label{eq:4DBICut}
\bubCutStates{$-$}{$-$}{$+$}{$+$} &= \sum_{h_i\in \text{states}} \mathcal{M}^{\text{BI}}(1^-, 2^-, l_2^{h_2},l_1^{h_1})\mathcal{M}^{\text{BI}}(-l_1^{\bar{h}_1},-l_2^{\bar{h}_2},3^+,4^+)
\\
&= \langle12\rangle^2[l_2 l_1 ]^2 \langle l_1 l_2\rangle^2 [34]^2
\\
&= [(l_1+l_2)^2]^2 \langle12\rangle^2 [34]^2
\\
&= s_{12}^2\langle12\rangle^2 [34]^2\,,
\end{align}
where in the last step we have made the replacement $(l_1+l_2) = -(k_1+k_2)$ by momentum conservation. Determining the leading behavior of the 1-loop Born-Infled amplitude thus amounts to specifying all possible 4D cuts.

@@

As we can see from the diagrams that contribute at two-loops four-point in \eqn{eq:2loopBasis}, even-point theories can be fully constructed from convolutions of one-loop integrals. As described earlier, we call such integrals {recursively one-loop} since we can iteratively apply one-loop unitarity and tensor reduction methods in order to recover the full multi-loop structure. In the next two sections, we first illustrate the recursive behavior of convolution integrals, and then use their properties to develop multi-loop tensor reduction methods that we use throughout the paper.

@@

    \caption{Graphical depiction of the the ostrich integral. Every exposed internal edge represents a propagator in the integrand.}

@@

\eta_{\mu_1\mu_2}I^{\mu_1\dots \mu_n}_{3,x}= \int \frac{d^D l}{(2\pi)^D}\frac{l^{\mu_2}\cdots l^{\mu_n} }{(l+K_{1})^{2x}(l+K_{12})^2} = 0\,.

@@

By iteratively applying Jacboi relations on internal edges, any color structure can be expressed completely in terms of functional relabelings of these two graphs. Concretely, these cubic graphs come dressed with the following color factors written in terms of group theory structure constants:

@@

We write the two-loop amplitude in this suggestive form to emphasize that the logarithmic divergence present at one-loop in \eqn{eq:1loopCP1} appears to exponentiate! More concretely, we have demonstrated through explicit calculation that through two-loop order, the leading divergence of the full color dressed amplitude goes as follows:

>>>>

The remarkable simplicity of one-loop amplitudes is even more dramatic in a fixed spacetime dimension. In $D=4$, the kinematic projection to a basis of scalar integrals saturates at the box integral due to the appearance of Gram determinants. Roughly speaking, $N$-gon integral coefficients of \eqn{TenRed1loop} come dressed with a factor of $\mathcal{G}_N$,
\begin{equation}
\mathcal{G}_N = \text{det} (k_i \cdot k_j)\,,
\end{equation}
where $k_i$ are the external momenta flowing into the $N$-gon integral. In fixed spacetime dimension, $D$, the momenta are $D$-component vectors and thus the Lorentz product matrix $(k_i \cdot k_j)$ must have a null space for $N>D$. As a result, $C_M \neq 0$, only when $M\leq D+1$. Furthermore, in $D=4-2\epsilon$ dimensions, any scalar pentagon integral can be rewritten up to $\mathcal{O}(\epsilon)$ in terms of five pinched scalar box integrals \cite{Bern:1993kr}. All the information of one-loop 4D amplitudes can therefore be completely determined by evaluating up to box integrals, regardless of the multiplicity. We will demonstrate in the next section how this can be used in 4D integral construction.

@@

\subsection{Even-point Effective Field Theories}
\label{subsec:EPEFTReview}
In this section we provide a catalog of the even-point effective field theories studied throughout the paper. 
Some of the theories we consider, like Volkov-Akulov and supersymmetric extensions of Born-Infeld, have natural formulations in terms of vierbeins that depend on fermionic degrees of freedom. For such theories, it is useful to distinguish between the coordinate independent actions, $S$, and the coordinate-dependent Lagrangian densities, $\mathcal{L}$,
\begin{equation}
S= \int d^D x \,\mathcal{L}\,.
\end{equation}
\paragraph{Nonlinear Sigma Model}
The simplest effective field theory that we will consider is the {nonlinear sigma model} (NLSM). The chiral NLSM Lagrangian density with diagonal symmetry group $G=SU(N_c)$ can be expressed in terms of the chiral current, $j_\mu = U^\dagger \partial_\mu U$, as follows,
\begin{equation}\label{eq:NLSMLag}
\mathcal{L}^{\text{NLSM}}= \frac{1}{2}\text{tr}[(\partial_\mu U)^\dagger (\partial ^\mu U)] = \frac{1}{2}\text{tr}\left[\frac{(\partial_\mu\pi )( \partial^\mu \pi) }{(1-f_\pi^{-2}\pi^2)^2}\right]\,,
\end{equation}
where the trace is over the color indices of the gauge group, $\pi \equiv \pi^a T^a$. On the right hand side we have applied the Cayley parameterization to the $SU(N_c)$ group elements, $U= \frac{1+\pi/f_\pi}{1-\pi/f_\pi}$, where $f_\pi$ is the dimensionful pion decay width. By construction, this model has a left-right global symmetry, $G\times G$, that is nonlinearly realized, and a linearly realized diagonal subgroup, $G$. This theory is the unique two-derivative EFT that is invariant under constant shifts of the pion field in color space:
\begin{equation}\label{eq:shiftSym}
\pi^a \rightarrow \pi^a + c^a\,.
\end{equation}
We could also consider higher derivative generalizations of the NLSM that are invariant under the shift symmetry of \eqn{eq:shiftSym}.  Such  EFTs correspond to the so-called chiral limit of {chiral perturbation theory}~\cite{Weinberg:1978kz,Gasser:1983yg,Kampf:2006yf,Kampf:2013vha}, describing massless pion scattering below the chiral symmetry breaking scale of QCD:
\begin{equation}\label{eq:XPTLag}
\mathcal{L}^{\chi\text{PT}}=\frac{1}{2}\text{tr}[(\partial_\mu U)^\dagger (\partial ^\mu U)] + \frac{\beta_1}{f_\pi^2} \text{tr}[(\partial_\mu U)^\dagger (\partial ^\mu U)]^2 + \frac{\beta_2}{f_\pi^2} \text{tr}[(\partial_\mu U)^\dagger( \partial_\nu U)]\text{tr}[(\partial^\mu U)^\dagger( \partial^\nu U)] +\cdots
\end{equation}
By construction, each of the higher derivative operators appearing a bove are invariant under the shift symmetry of \eqn{eq:shiftSym}. These higher derivative generalizations of NLSM have been studied both in the context of the $S$-matrix bootstrap \cite{Manohar:2008tc,Bellazzini:2016xrt,Guerrieri:2020bto}, and soft theorem bootstraps of Refs. \cite{Cheung:2014dqa,Cheung:2015ota,Cheung:2016drk,Cheung:2018oki,Low:2019ynd}. Given the relevance of \eqn{eq:XPTLag} in Standard Model pion scattering, the amplitudes for $G= SU(2)$ isospin pions have been computed through two-loop order \cite{Bijnens:1995yn,Bijnens:1997vq,Girlanda:1997ed}. In this work, we add to the literature by computing the two-loop amplitudes of \eqn{eq:NLSMLag} for arbitrary $SU(N_c)$ color structure. We then use these amplitudes to compute $\mathcal{N}=4$ DBIVA amplitudes via the double copy in \sect{sec:DBIvDC}.

@@

\paragraph{Volkov-Akulov Fermions and Supersymmetry}
Now we will describe the supersymmetric extensions of the Born-Infeld action given above in \eqn{biLag}. First we start by defining the theory of shift symmetric fermions first written down by Volkov, Akulov (VA)~\cite{Volkov:1973ix}. The easiest way to define this shift symmetric VA theory is by constructing a volume form in the vierbein frame out of manifestly supersymmetric 1-forms. That is, define a 1-form, $ \omega^m = e^m_\mu dx^\mu$, where the vierbeins (frame fields), $e^m_\mu$, are expressed in terms of the fermion fields, $\lambda$, as follows:
 \begin{equation}
 e^m_\mu = \delta^m_\mu + i \bar{\lambda} \Gamma^m\! \stackrel{\leftrightarrow}{\partial_\mu}\!  \lambda = \delta^m_\mu + i (\bar{\lambda} \Gamma^m \partial_\mu  \lambda - \partial_\mu \bar{\lambda} \Gamma^m   \lambda)\,.
 \end{equation}
 This frame field is manifestly invariant under the superspace shift:
\begin{equation}
\delta \lambda \rightarrow \eta \qquad  \delta \bar{\lambda} \rightarrow \bar{\eta} \qquad  \delta x^\mu \rightarrow x^\mu - i (\bar{\lambda} \,\Gamma^\mu \eta- i \bar{\eta} \,\Gamma^\mu \lambda)\,.
\end{equation}
 The Volkov-Akulov action in the vierbein frame is then simply the fermionic world-volume integral over the $D$-form \cite{Kallosh:1997aw},
 \begin{equation}
 S^{\text{VA}} = \int \omega^1 \wedge \omega^2 \wedge \cdots \wedge \omega^D\,.
 \end{equation}
 This is not dissimilar from the construction of the pure scalar DBI action above. Likewise, we can write the Lagrangian in the spacetime coordinate frame in parallel to \eqn{biLag} and \eqn{DBI}. Transforming back into spacetime coordinates yields the following Lagrangian density for self interaction VA fermions upto four-point interactions:
\begin{equation}
 \label{vaLag}
  \alpha'^{\,2}\mathcal{L}^{\text{VA}} = \text{det}(e^m_\mu) =  i\bar{\lambda}\!\stackrel{\leftrightarrow}{\slashed{\partial}}\!\lambda+ \frac{1}{2}(\bar{\lambda}\Gamma^\mu \partial^\nu \lambda)(\bar{\lambda}\Gamma_\mu \partial_\nu \lambda)+\mathcal{O}(\partial^4\lambda^6) 
  \,.
\end{equation}
Furthermore, a convenient choice of field redefinition was proposed by Komargodski and Seiberg (KS) \cite{Komargodski:2009rz,Kuzenko:2010ef}, which defines a new fermion field, $\psi = \psi(\lambda,\bar{\lambda})$, in a way that is perturbatively equivalent to the VA fermions. This construction is equivalent to a theory of nilpotent chiral superfields \cite{Rocek:1978nb,Casalbuoni:1988xh,Ferrara:2014kva}, relevant for the construction of $\alpha$-attractor models of inflation \cite{Kallosh:2013yoa}. Applying the KS nonlinear field redefinition yields the considerably simpler Lagrangian for VA theory:
 \begin{equation}
   \alpha'^{\,2}\mathcal{L}^{\text{VA}} = i\bar{\psi}\slashed{\partial}\psi +\frac{1}{2} \bar{\psi}^2 \partial^2 \psi^2 + \frac{1}{4}\psi^2\bar{\psi}^2 \partial^2 \psi^2\partial^2 \bar{\psi}^2\,.
 \end{equation}
With the world-volume formulation of VA theory in hand, we can also trivially construct the action for maximal $\mathcal{N}=1$ supersymmetric Born-Infeld theory in $D=10$ \cite{Tseytlin:1999dj}. We simply replace the flat space background of \eqn{biLag} with the fermionic background of \eqn{vaLag},
\begin{equation}
 \label{DBIVALag}
S^{\text{DBIVA}}_{\mathcal{N}=1} = \int \omega^1 \wedge\omega^2 \wedge \cdots \wedge \omega^{10} \sqrt{\text{det}(\eta_{mn}+ \alpha F_{mn})}\,.
\end{equation}
It is worth noting that the fermion-vector interaction is introduced via the {non-minimal} coupling between the field strength and the fermionic spacetime metric \cite{Bergshoeff:2013pia}. This differs from the linearly realized supersymmetry of super-Yang-Mills theory, which minimally couples the vector mode to the chiral fermion. This non-minimal coupling is a requirement of supersymmtry due to the even-point nature of abelian vector theories like Born-Infeld.

@@

\subsection{On-shell Unitarity methods}\label{sec:genU}
Above we have reviewed the known behavior of one-loop integral bases in $D=4$ spacetime dimensions. In this section we will briefly demonstrate why traditional 4D unitarity methods are so powerful -- and why we must go beyond them to accommodate the multi-loop calculations of interest here. 
\paragraph{Integrands from 4D Cuts}
Applying one-loop integral reduction in 4D tells us that any one-loop massless amplitude, for which tadpoles are scaleless, can be expressed in terms of box, triangle, and bubble diagrams, with each integral weighted by dimensionally-dependent functions of kinematics:
\begin{equation}\label{eq:4DUn}
\mathcal{A}^{\text{1-loop}}= \sum_{N=2}^4 C_N(D)I_N^D = \sum_{N=2}^4 C_N(4)I_N^D+ \mathcal{R}\,,
\end{equation}
where the scalar basis-integrals, $I_N^D$, are evaluated in $D=4-2\epsilon$, and $\mathcal{R}$ is a rational term that emerges from series expanding around $\epsilon=(4-D)/2$. To construct integral coefficients, $C_N$, we make use of the Optical Theorem for the $S$-matrix, which states that,
\begin{equation}
S^\dagger S = \mathbb{1} \quad \Rightarrow \quad 2\,\text{Im}(T) = T^\dagger T\,,
\end{equation}
where $S=\mathbb{1}+iT$. Applying this relation to \eqn{eq:4DUn} allows us to identify the imaginary part of $\mathcal{A}^{\text{1-loop}}$ with the imaginary parts of each basis integral weighted by the respective 4D coefficient, $C_N(4)$:
\begin{equation}\label{eq:UnMeth1loop}
\text{Im}(\mathcal{A}^{\text{1-loop}}) = \sum_{N=2}^4 C_N(4)\, \text{Im}(I_N^D)\,.
\end{equation}
Using the optical theorem to constrain scattering amplitudes is known as the \textit{unitarity method} \cite{UnitarityMethod, BDKUniarityReview}. In order to isolate each $C_N(4)$ appearing in \eqn{eq:4DUn}, we must place particular internal legs on-shell. That is, for each internal loop propagator that we wish to take on-shell, $1/P_i^2$, we make the following replacement inside the integral:
\begin{equation}
\frac{i}{P_i^2 + i\varepsilon} \rightarrow 2\pi \delta^{(+)}(P_i^2+i\varepsilon)\,.
\end{equation}
By replacing internal propagators with positive-root delta functions, one can isolate the imaginary parts needed to determine the integral coefficients. This approach of taking individual legs on-shell is known as the method of \textit{generalized unitarity} -- and it can be systematically applied to construct all one-loop amplitudes directly from 4D unitarity cuts \cite{Forde:2007mi}. Using generalized unitarity, the integral coefficients correspond to the following set of iterated cuts:
\begin{align}
C^{(P_iP_jP_k)}_4 &\sim \boxCut\,,
\\
C^{(P_iP_j)}_3 &\sim \triCut - \sum_{k} C^{(P_iP_jP_k)}_4\,,
\\
C^{(P_i)}_2 &\sim \bubCut- \sum_{j} C_3^{(P_iP_j)} - \sum_{j,k} C^{(P_iP_jP_k)}_4\,,
\end{align}
where exposed legs are summed over all internal states crossing the cut, and the labels $P_i$ are external momentum scales flowing into the specified cuts. The subtractions account for the overlapping information in the respective $N$-gon cuts. By momentum conservation, the momentum flow in the unlabeled vertex is completely determined by the other $N-1$ vertices. As a concrete example, consider the 4D cut construction of the bubble coefficient for pure Born-Infeld theory:
\begin{align}\label{eq:4DBICut}
\bubCutStates{$-$}{$-$}{$+$}{$+$} &= \sum_{h_i\in \text{states}} \mathcal{M}^{\text{BI}}(1^-, 2^-, l_2^{h_2},l_1^{h_1})\mathcal{M}^{\text{BI}}(-l_1^{\bar{h}_1},-l_2^{\bar{h}_2},3^+,4^+)
\\
&= \langle12\rangle^2[l_2 l_1 ]^2 \langle l_1 l_2\rangle^2 [34]^2
\\
&= [(l_1+l_2)^2]^2 \langle12\rangle^2 [34]^2
\\
&= s_{12}^2\langle12\rangle^2 [34]^2\,,
\end{align}
where in the last step we have made the replacement $(l_1+l_2) = -(k_1+k_2)$ by momentum conservation. Determining the leading behavior of the 1-loop Born-Infled amplitude thus amounts to specifying all possible 4D cuts.

@@

As we can see from the diagrams that contribute in \eqn{eq:2loopBasis}, the scattering amplitudes for even-point theories at two-loop four-point can be fully constructed from convolutions of one-loop integrals. As mentioned earlier, we call such integrals {recursively one-loop} since we can iteratively apply one-loop unitarity and tensor reduction methods in order to recover the full multi-loop structure. We note that at general multiplicity and at higher loop order, EMU cut construction will produce integrals outside the space of recursively one-loop integrals. While it would absolutely be worthwhile to determine the EMU integral basis in generality, in this paper we focus on the nice example of two-loop four-point. To this end, in the next two sections we will illustrate the recursive behavior of one-loop convolution integrals, and then use their properties to develop multi-loop tensor reduction methods for the two-loop four-point amplitudes considered in this work.

@@

    \caption{Graphical depiction of the the ostrich integral. Every exposed internal edge represents a propagator in the integrand. Throughout the text, ostrich integrals will be drawn with an asymmetry between the ``head'' and the ``tail'' of the ostrich, but we emphasize that the integral itself is functionally symmetric under exchange of the top two external edges.}

@@

\eta_{\mu_1\mu_2}I^{\mu_1\dots \mu_n}_{3,x}= \int \frac{d^D l}{(2\pi)^D}\frac{l^{\mu_3}\cdots l^{\mu_n} }{(l+K_{1})^{2x}(l+K_{12})^2} = 0\,.

@@

By iteratively applying Jacobi relations on internal edges, any color structure can be expressed completely in terms of functional relabelings of these two graphs. Concretely, these cubic graphs come dressed with the following color factors written in terms of group theory structure constants:

@@

We write the two-loop amplitude in this suggestive form to emphasize that the logarithmic divergence present at one-loop in \eqn{eq:1loopCP1} appears to exponentiate! More concretely, we have demonstrated through explicit calculation that through two-loop, the leading divergence of the full color dressed amplitude at each loop order goes as,
